### 2021-04-24 notes/penetration_argument/penetration_argument.tex
adds zeroth order penetration argument doc

\documentclass[12pt,preprint]{article}
\usepackage{natbib}
\bibliographystyle{plain}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}

\usepackage{pb-diagram}

\parindent = 0.0 in
\parskip = 0.15 in

\newcommand\Beq{\begin{align}} 
\newcommand\Eeq{\end{align}}

\newcommand\Bfig{\begin{figure}} 
\newcommand\Efig{\end{figure}}

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

\newcommand\eps{\varepsilon}
\renewcommand\L {\mathcal{L}}

\newcommand{\n}{\\ \nonumber \\ }
\newcommand{\nn}{\nonumber}
\newcommand{\nnn}{\\ \nonumber \\ \nonumber}

\newcommand\ie{\textit{i.e.},~}
\newcommand\eg{\textit{e.g.},~}
\newcommand{\omicron}{o}

\newcommand{\pd}[1]{\partial_{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\M}[1]{\mathbf{#1}}
\renewcommand{\dot}{\vec{\cdot}}
\newcommand{\grad}{\vec{\nabla}}
\newcommand{\cross}{\vec{\times}}
\newcommand{\laplacian}{\nabla^2}

\newcommand{\sump}[2]{\sideset{}{'}\sum_{{#1}=0}^{#2}}

\newcommand{\eq}[1]{eq.~(\ref{#1})}
\newcommand{\eqs}[2]{eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\eqss}[2]{eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\Eqss}[2]{Eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\figs}[2]{Figs.~(\ref{#1})~\&~(\ref{#2})}

\newcommand{\tbl}[1]{Table~(#1)}

\begin{document}

%\section{Nondimensional equations}
%The nondimensional system of equations we solve is
%\begin{align}
%&\grad\dot\vec{u} = 0 \\
%&\partial_t \vec{u} + \vec{u}\dot\grad\vec{u} = -\grad \varpi + T \hat{z} + \rm{Re}_0^{-1}\grad^2 \vec{u} \\
%&\partial_t T + \vec{u}\dot\grad T + w \grad_{\rm{ad}} = \rm{Pe}_0^{-1}\grad^2 T' + \grad\dot[k \grad \overline{T}] + Q
%\end{align}

\section{Convective Penetration argument}
First some definitions:
\begin{enumerate}
\item $z=L_s$ is the the top of the convection zone according to the Schwarzschild criterion; it's the height where $\grad_{\rm{ad}} = \grad_{\rm{rad}}$.
\item $L_{\rm{cz}}$ is the top of the convection zone; it's roughly the top of the region where convection flattens $\grad \rightarrow \grad_{\rm{ad}}$.
We generally get $L_{\rm{cz}} > L_s$.
\item The penetration depth is $\delta_p = L_{\rm{cz}} - L_s$.
\item The flux carried by convection for $z < L_s$ is $F_{\rm{conv,cz}} = Q \delta_H$, where $Q$ is the magnitude of the internal heating and $\delta_H$ is the depth of the heating layer.
\item $\overline{w}$ is the vertical profile of the characteristic (vertical) convective velocity, which is a constant $w_{\rm{cz}}$ for $z \leq L_{\rm{cz}}$.
\item Similarly, $\overline{\delta T}$ is the vertical profile of the characteristic temperature perturbation.
\end{enumerate}

And some key assumptions.
\begin{enumerate}
\item Convection flattens $\grad \rightarrow \grad_{\rm{ad}}$ for $z \leq L_{\rm{cz}}$ (baked into our definitions).
\item We assume a system in thermal equilibrium, at least in (adiabatic) convection zone with $z \leq L_{\rm{cz}}$.
Therefore $F_{\rm{conv}}(z) = F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)$, where $F$ is a flux and $F_{\rm{rad, ad}}$ is the radiative flux along the adiabatic gradient.
\item We assume $F_{\rm{conv}} = \overline{w T} \approx \overline{w}\overline{\delta T}$.
Combined with our previous assumption, we get
\begin{equation}
\overline{\delta T} \approx \frac{F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)}{\overline{w}}.
\end{equation}
\end{enumerate}
As we increase in height, the conductivity and thus $F_{\rm{rad, ad}}(z)$ also increases.
This means that $\overline{\delta T}$ has the opposite sign of $\overline{w}$ for $z > L_s$.
We presume that buoyancy breaking is the dominant mechanism which brings convective motions to a stop in this adiabatic layer,
\begin{equation}
\frac{d \overline{w}}{dt} = \overline{\delta T} = \frac{F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)}{\overline{w}}.
\end{equation}
If we multiply both sides by $\overline{w}$ and absorb it into the the derivative, then apply the chain rule with $d/dt = d/dz (dz/dt) = d/dz \overline{w}$, we retrieve,
\begin{equation}
\frac{1}{2} d\overline{w}^3 = \left[F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)\right]dz.
\end{equation}
We now replace $F_{\rm{rad, ad}}(z) = k_0(z) \grad_{\rm{ad}}$, and we assume that $k(z)$ instantaneously jumps from a low value $k_{\rm{cz}}$ to a high value $k_{\rm{rz}}$ at $z = L_s$.
Per this assumption, the convective flux is a negative constant for $z \geq L_s$.
We integrate from $z = L_s$ with $\overline{w} = w_{\rm{cz}}$ to $z = L_{s}$ with $\overline{w} = 0$, and get
\begin{equation}
-\frac{w_{\rm{cz}}^3}{2} = F_{\rm{conv,p}} \delta_p.
\end{equation}
By definition, the convective flux in the penetrative layer with $z > L_s$ is related to $\mathcal{P}$,
\begin{equation}
F_{\rm{conv,t}} = -\frac{F_{\rm{conv,cz}}}{\mathcal{P}},
\end{equation}
so we retrieve
\begin{equation}
\delta_p = \frac{w_{\rm{cz}}^3 \mathcal{P}}{2 F_{\rm{conv, cz}}},
\end{equation}
and the penetration depth scales with the cube of the velocity and linearly with $\mathcal{P}$.
This prediction does not exactly line up with the simulation results (this overpredicts the magnitude of $\delta_p$).
The important thing we need to determine is if it scales with $w_{\rm{cz}}$, $\mathcal{P}$, and $F_{\rm{conv,cz}}$ in the appropriate way.
So far, it seems to scale linearly with $\mathcal{P}$ at fixed other parameters.

\bibliography{biblio.bib}
\end{document}