### 2021-04-28 notes/penetration_argument/penetration_argument.tex
updated dimensional penetration argument

\documentclass[12pt,preprint]{article}
\usepackage{natbib}
\bibliographystyle{plain}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}

\usepackage{pb-diagram}

\parindent = 0.0 in
\parskip = 0.15 in

\newcommand\Beq{\begin{align}} 
\newcommand\Eeq{\end{align}}

\newcommand\Bfig{\begin{figure}} 
\newcommand\Efig{\end{figure}}

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

\newcommand\eps{\varepsilon}
\renewcommand\L {\mathcal{L}}

\newcommand{\n}{\\ \nonumber \\ }
\newcommand{\nn}{\nonumber}
\newcommand{\nnn}{\\ \nonumber \\ \nonumber}

\newcommand\ie{\textit{i.e.},~}
\newcommand\eg{\textit{e.g.},~}
\newcommand{\omicron}{o}

\newcommand{\pd}[1]{\partial_{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\M}[1]{\mathbf{#1}}
\renewcommand{\dot}{\vec{\cdot}}
\newcommand{\grad}{\vec{\nabla}}
\newcommand{\cross}{\vec{\times}}
\newcommand{\laplacian}{\nabla^2}

\newcommand{\sump}[2]{\sideset{}{'}\sum_{{#1}=0}^{#2}}

\newcommand{\eq}[1]{eq.~(\ref{#1})}
\newcommand{\eqs}[2]{eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\eqss}[2]{eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\Eqss}[2]{Eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\figs}[2]{Figs.~(\ref{#1})~\&~(\ref{#2})}

\newcommand{\tbl}[1]{Table~(#1)}

\begin{document}

\section{Dimensional equations}
The dimensional Boussinesq equations are
\begin{align}
&\grad\dot\vec{u} = 0 \label{eqn:continuity} \\
&\partial_t \vec{u} + \vec{u}\dot\grad\vec{u} = -\frac{1}{\rho_0}\grad p + \frac{\rho'}{\rho_0}\vec{g} + \nu\grad^2 \vec{u} \label{eqn:momentum} \\
&\partial_t T + \vec{u}\dot\grad T + w \grad_{\rm{ad}} = \chi\grad^2 T' + \grad\dot[k \grad \overline{T}] + Q \label{eqn:temperature} \\
&\frac{\rho'}{\rho_0} = -|\alpha| T \label{eqn:boussinesq}
\end{align}
Where $\rho$ is the density, $T$ is the temperature, $\vec{g}$ is the gravitational acceleration, $\alpha$ is the coefficient of thermal expansion, $\nu$ and $\chi$ are the viscous and thermal diffusivity, $k$ is a radiative diffusivity, and $Q$ is a heating term.
We have baked in an assumption that $\alpha < 0$ to make sign conventions more straightforward after substituting Eqn.~\ref{eqn:boussinesq} into Eqn.~\ref{eqn:momentum}.

\section{Convective Penetration argument}
First some definitions:
\begin{enumerate}
\item $z=L_s$ is the the top of the convection zone according to the Schwarzschild criterion; it's the height where $\grad_{\rm{ad}} = \grad_{\rm{rad}}$.
\item $L_{\rm{cz}}$ is the top of the convection zone; it's roughly the top of the region where convection flattens $\grad \rightarrow \grad_{\rm{ad}}$.
We generally get $L_{\rm{cz}} > L_s$.
\item The penetration depth is $\delta_p = L_{\rm{cz}} - L_s$.
\item The flux carried by convection for $z < L_s$ is $F_{\rm{conv,cz}} = Q \delta_H$, where $Q$ is the magnitude of the internal heating and $\delta_H$ is the depth of the heating layer.
\item $\overline{w}$ is the vertical profile of the characteristic (vertical) convective velocity, which is a constant $w_{\rm{cz}}$ for $z \leq L_{\rm{cz}}$.
\item Similarly, $\overline{\delta T}$ is the vertical profile of the characteristic temperature perturbation.
\end{enumerate}

And some key assumptions.
\begin{enumerate}
\item Convection flattens $\grad \rightarrow \grad_{\rm{ad}}$ for $z \leq L_{\rm{cz}}$ (baked into our definitions).
\item We assume a system in thermal equilibrium, at least in (adiabatic) convection zone with $z \leq L_{\rm{cz}}$.
Therefore $F_{\rm{conv}}(z) = F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)$, where $F$ is a flux and $F_{\rm{rad, ad}}$ is the radiative flux along the adiabatic gradient.
\item We assume $F_{\rm{conv}} = \overline{w T} \approx \overline{w}\overline{\delta T}$.
Combined with our previous assumption, we get
\begin{equation}
\overline{\delta T} \approx \frac{F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)}{\overline{w}}.
\label{eqn:dT_assumption}
\end{equation}
\end{enumerate}
As we increase in height, the conductivity and thus $F_{\rm{rad, ad}}(z)$ also increases.
This means that $\overline{\delta T}$ has the opposite sign of $\overline{w}$ for $z > L_s$.

We presume that buoyancy breaking is the dominant mechanism which brings convective motions to a stop in this adiabatic layer.
In other words, we assume that, if we drop the nonlinear, pressure, and viscous terms from Eqn.~\ref{eqn:momentum}, we describe the dynamics reasonably well,
\begin{equation}
\frac{d \overline{w}}{dt} = \frac{\overline{\rho'}}{\rho_0}(-g\hat{z}) = \alpha g\, \overline{\delta T} = \alpha g \, \frac{F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)}{\overline{w}}.
\end{equation}
If we multiply both sides by $\overline{w}$ and absorb it into the the derivative, then apply the chain rule with $d/dt = d/dz (dz/dt) = d/dz \overline{w}$, we retrieve,
\begin{equation}
\frac{1}{2} d\overline{w}^3 = \alpha g\, \left[F_{\rm{tot}}(z) - F_{\rm{rad, ad}}(z)\right]dz.
\end{equation}
We now replace $F_{\rm{rad, ad}}(z) = k_0(z) \grad_{\rm{ad}}$, and we assume that $k(z)$ instantaneously jumps from a low value $k_{\rm{cz}}$ to a high value $k_{\rm{rz}}$ at $z = L_s$.
Per this assumption, the convective flux is a negative constant for $z \geq L_s$.
We integrate from $z = L_s$ with $\overline{w} = w_{\rm{cz}}$ to $z = L_{s}$ with $\overline{w} = 0$, and get
\begin{equation}
-\frac{w_{\rm{cz}}^3}{2} = \alpha g\, F_{\rm{conv,p}}\,  \delta_p.
\end{equation}
By definition, the convective flux in the penetrative layer with $z > L_s$ is related to $\mathcal{P}$,
\begin{equation}
F_{\rm{conv,p}} = -\frac{F_{\rm{conv,cz}}}{\mathcal{P}},
\end{equation}
so we retrieve
\begin{equation}
\delta_p = \frac{w_{\rm{cz}}^3}{2\alpha g\, F_{\rm{conv, cz}}}  \mathcal{P},
\end{equation}
and the penetration depth scales with the cube of the velocity and linearly with $\mathcal{P}$.
This prediction does not exactly line up with the simulation results (this overpredicts the magnitude of $\delta_p$).
The important thing we need to determine is if it scales with $w_{\rm{cz}}$, $\mathcal{P}$, and $F_{\rm{conv,cz}}$ in the appropriate way.
So far, it seems to scale linearly with $\mathcal{P}$ at fixed other parameters.

Finally, we note that for an ideal gas with $P = \mathcal{R} \rho T$,
\begin{equation}
\alpha = \frac{\partial \ln\rho}{\partial T} = \frac{1}{\rho}\frac{\partial\rho}{\partial T} \approx - T^{-1}.
\end{equation}
If we had assumed that the convective flux had the form $\rho c_p w \delta T$ during Eqn.~\ref{eqn:dT_assumption}, we would have retrieved
\begin{equation}
\delta_p = \frac{\rho\, c_p\, w_{\rm{cz}}^3}{2\alpha g\, F_{\rm{conv, cz}}}  \mathcal{P}
\qquad \Rightarrow \qquad
\frac{\delta_p}{H_P} = \frac{\rho\, w_{\rm{cz}}^3}{2 F_{\rm{conv, cz}}}  \mathcal{P},
\end{equation}
with $H_P = c_P T / g$ the rough pressure scale height.

\bibliography{biblio.bib}
\end{document}